Initial spatial conditions in simulation models: the missing leg of sensitivity analyses?

### 2017-11-03 paper.tex
Update on Overleaf.

In order to test for the influence of initial spatial conditions on model outputs, we use a systematic method to compare phase diagrams. We define a phase diagram as the vector of model outputs considered as a function of model parameters. Technically, because of stochasticity, we represent the output of the model for a given combination of parameter values $\overrightarrow{P}$ as the median of the final values of an output indicator $O$ obtained for the replications of the model initialized with $\overrightarrow{P}$.\\

%Comparing phase diagrams is useful because it allows us to compare the 
Indeed, we have as many phase diagrams than we have spatial grids, what makes a qualitative visual comparison not realistic (with around 50 different spatial configurations in our applications). A solution is to use systematic quantitative procedures. To our knowledge there exist no well established method to compare phase diagrams in agent-based modeling and geosimulation literature. We propose to use a relative distance between two phase diagram corresponding to the share of between-diagrams variability relative to their internal variability, given formally in the case of a one-dimensional phase diagram by

@@

The last methodological point which we need to emphasis is the relation between the workflow we introduce and model exploration workflows. The ideas of multi-modeling and extensive model exploration are nothing from new as ~\cite{openshaw1983data} already advocated for ``model-crunching'', but their effective use only begins to emerge thanks to the apparition of new methods and tools together with an explosion of computation capabilities. The model exploration platform OpenMole~\citep{reuillon2013openmole} allows to embed any model as a blackbox, write modulable exploration workflow using advanced methodologies such as genetic algorithms and distribute transparently the computation on large scale computation infrastructures such as clusters or computation grids. In our case, the workflow tool is a powerful way to embed both the sensitivity analysis and the meta-sensitivity analysis, and allow to couple any generator with any model in a straightforward way as soon as the model can take its spatial configuration as input or from an input file. In this paper, we use the OpenMole platform for spatial environment and model coupling, placing ourselves in the renewed framework of multi-modelling claimed by \citet{cottineau2015modular}.
%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial generator of density grids}

>>>>

In order to test for the influence of initial spatial conditions on model outputs, we use a systematic method to compare {\bf phase diagrams}. We define a phase diagram as the vector of model outputs considered as a function of model parameters. Technically, because of stochasticity, we represent the output of the model for a given combination of parameter values $\vec{P}$ as the median of the final values of an output indicator $O$ obtained for the replications of the model initialized with $\vec{P}$.\\

%Comparing phase diagrams is useful because it allows us to compare the 
We have as many phase diagrams than we have spatial grids, which makes a qualitative visual comparison not realistic (with around 50 different spatial configurations for each model experiment). A solution is to use systematic quantitative procedures to compare them. To our knowledge there exist no single well established method to compare phase diagrams in the agent-based modelling and geosimulation literature. We propose to define the relative distance $d_r$ between two phase diagrams $\vec{\alpha_1}$ and $\vec{\alpha_2}$ as the share of between-diagrams variability relative to their internal variability, given formally in the case of a one-dimensional phase diagram by

@@

The last methodological point which we need to emphasize is the relation between the workflow we introduce and model exploration workflows. The ideas of multi-modeling and extensive model exploration are nothing from new as ~\cite{openshaw1983data} already advocated for ``model-crunching'', but their effective use only begins to emerge thanks to the apparition of new methods and tools together with an explosion of computation capabilities. The model exploration platform OpenMole~\citep{reuillon2013openmole} allows to embed any model as a blackbox, write modulable exploration workflow using advanced methodologies such as genetic algorithms and distribute transparently the computation on large scale computation infrastructures such as clusters or computation grids. In our case, the workflow tool is a powerful way to embed both the sensitivity analysis and the meta-sensitivity analysis, and allow to couple any generator with any model in a straightforward way as soon as the model can take its spatial configuration as input or from an input file. In this paper, we use the OpenMole platform for spatial environment and model coupling, placing ourselves in the renewed framework of multi-modelling claimed by \citet{cottineau2015modular}.
%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial generator of density grids}
